---
title: "Spectral pre-processing recipes"
author:
 - name: Leonardo Ramirez-Lopez and Claudio Orellano
   email: ramirez-lopez.l@buchi.com
   affiliation: Data Science Department, BUCHILabortechnik AG, Flawil, Switzerland
date: today
clean: true
bibliography: ["proximetricsR.bib"]
biblio-style: "apalike"
link-citations: true
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-overflow: wrap
    smooth-scroll: true
    html-math-method: mathjax
vignette: >
  %\VignetteIndexEntry{Spectral pre-processing recipes}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{quarto::html}
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  eval = requireNamespace("proximetricsR", quietly = TRUE)
)

# Disable ANSI colours for vignette rendering
options(cli.num_colors = 1)
Sys.setenv("RSTUDIO" = "")
Sys.setenv("POSITRON" = "")
old_options <- options(digits = 3)
```

# Overview

Spectral preprocessing is a critical step in near-infrared (NIR) calibration workflows. Raw spectral data often contains systematic variations, noise, and artifacts that can obscure the true relationship between spectra and reference properties. The `proximetricsR` package provides a flexible, composable system for building and applying preprocessing pipelines.

## Key concepts

- **Preprocessing constructors** (`prep_*()` functions): Create specifications for individual preprocessing steps
- **Preprocessing recipes** (`preprocess_recipe()`): Assemble multiple constructors into an ordered sequence
- **Recipe execution** (`process()`): Apply the recipe to spectral data matrices
- **Device compatibility**: Different devices (ProxiMate, ProxiScout) support different preprocessing algorithms

This separation of specification and execution enables reproducible, device-aware preprocessing pipelines that can be stored, shared, and applied consistently.

# Setup



```{r}
#| label: loadlib
#| results: 'hide'
#| include: False
knitr::opts_chunk$set(
  eval = requireNamespace("proximetricsR", quietly = TRUE)
)
library("proximetricsR")
```


```{r loadlib2, eval = FALSE}
library("proximetricsR")
```

```{r data, results = 'hide'}
data("NIRcannabis")
X <- NIRcannabis$spc
```

# Preprocessing constructors

The `prep_*()` functions create preprocessing step objects. Each constructor validates its parameters and encodes algorithm-specific information. The order in which constructors are passed to `preprocess_recipe()` defines the execution order.

## Resampling: `prep_resample()`

Resampling interpolates spectra to a new wavelength or wavenumber grid.

**ProxiMate mode** (user-defined grid):
```{r prep_resample_proximate}
prep_resample(grid = c(1001, 1700, 2))
```

**ProxiScout mode** (NeoSpectra standard grid):
```{r prep_resample_proxiscout}
prep_resample(grid = "proxiscout")
```

Resampling is often the first step to standardize wavelength grids across different instruments.

## Smoothing: `prep_smooth()`

Smoothing reduces high-frequency noise while preserving spectral features.

**Savitzky-Golay** (ProxiScout compatible):
```{r prep_smooth_sg}
prep_smooth(w = 11, p = 3, algorithm = "savitzky-golay")
```

**Moving average** (ProxiMate compatible):
```{r prep_smooth_ma}
prep_smooth(w = 7, algorithm = "moving-average")
```

- `w`: window size (must be a positive odd integer)
- `p`: polynomial order for Savitzky-Golay (must be < w)

## Standard Normal Variate: `prep_snv()`

SNV (Standard Normal Variate) normalizes each spectrum independently by centering and scaling:

$$SNV_i = \frac{x_i - \bar{x}_i}{s_i}$$

where $\bar{x}_i$ and $s_i$ are the mean and standard deviation of the $i$-th spectrum.

```{r prep_snv}
prep_snv()
```

SNV corrects for multiplicative effects (e.g., baseline offsets, path length variations) and is device-agnostic.

## Derivatives: `prep_derivative()`

Derivatives enhance spectral differences and reduce baseline effects.

**Savitzky-Golay** (ProxiScout):
```{r prep_derivative_sg}
prep_derivative(m = 1, w = 11, p = 3, algorithm = "savitzky-golay")
```

**Gap-Segment** (ProxiScout):
```{r prep_derivative_gap}
prep_derivative(m = 2, w = 9, p = 3, algorithm = "gap-segment")
```

**NIRWise PLUS compatible** (ProxiMate):
```{r prep_derivative_nwp}
prep_derivative(m = 1, w = 5, p = 11, algorithm = "nwp")
```

Parameters:
- `m`: derivative order (1 or 2)
- `w`: window/gap size (positive odd integer)
- `p`: polynomial order (Savitzky-Golay) or smoothing window (gap-segment, nwp)
- `algorithm`: choice of method

## Detrending: `prep_detrend()`

Detrending removes wavelength-dependent baseline effects by fitting and removing a polynomial trend (ProxiScout only):

```{r prep_detrend}
prep_detrend(p = 2)
```

- `p`: polynomial order (default 2)

For the full Barnes et al. (1989) procedure (SNV + detrending), chain `prep_snv()` before `prep_detrend()`.

## Reflectance/Absorbance conversion: `prep_transform()`

Convert between reflectance and absorbance using Beer's Law (ProxiScout only):

$$A = -\log_{10}(R)$$

```{r prep_transform}
prep_transform(to = "absorbance")
```

- `to`: target unit (`"absorbance"` or `"reflectance"`)

## Wavelength trimming: `prep_wav_trim()`

Retain only a specified wavelength band and/or remove constant-valued edge columns (ProxiScout only):

```{r prep_wav_trim}
prep_wav_trim(band = c(1000, 1800), trim_constant_edges = TRUE)
```

- `band`: wavelength range to retain (or `c()` to skip)
- `trim_constant_edges`: remove zero or constant-valued columns at edges

# Building preprocessing recipes

The `preprocess_recipe()` function combines constructors into an ordered pipeline. Order matters: preprocessing steps are applied in the order specified.

## Device compatibility

Different BUCHI devices support different preprocessing steps:

**ProxiMate** supports:
- `prep_resample()` with user-defined grids
- `prep_smooth()` with moving-average algorithm
- `prep_snv()`
- `prep_derivative()` with nwp algorithm

**ProxiScout** supports:
- `prep_resample()` with "proxiscout" grid
- `prep_smooth()` with savitzky-golay algorithm
- `prep_snv()`
- `prep_derivative()` with savitzky-golay or gap-segment algorithms
- `prep_detrend()`
- `prep_transform()`
- `prep_wav_trim()`

## Building recipes

**Single preprocessing step (SNV only):**

SNV is device-agnostic, so `device` is optional:

```{r recipe_snv}
recipe_snv <- preprocess_recipe(prep_snv())
recipe_snv
```

**Multiple steps (requires device):**

```{r recipe_proxiscout}
recipe_ps <- preprocess_recipe(
  prep_smooth(w = 7, p = 1, algorithm = "savitzky-golay"),
  prep_snv(),
  prep_derivative(m = 1, w = 5, p = 2, algorithm = "savitzky-golay"),
  device = "proxiscout"
)
recipe_ps
```

**ProxiMate-specific recipe:**

```{r recipe_proximate}
recipe_pm <- preprocess_recipe(
  prep_smooth(w = 7, algorithm = "moving-average"),
  prep_snv(),
  prep_derivative(m = 1, w = 5, p = 11, algorithm = "nwp"),
  device = "proximate"
)
recipe_pm
```

Recipes validate that all steps are compatible with the specified device and raise informative errors if not.

# Applying recipes with `process()`

The `process()` function executes a recipe on spectral data:

```{r process_snv}
X_snv <- process(X, recipe_snv)
dim(X_snv)
```

```{r process_proxiscout}
X_ps <- process(X, recipe_ps)
dim(X_ps)
```

The applied recipe is stored as an attribute and can be retrieved:

```{r process_inspect}
applied_recipe <- attr(X_ps, "preprocess_recipe")
applied_recipe
```

# Practical examples

## Example 1: ProxiMate workflow

A typical ProxiMate workflow for fat/protein prediction:

```{r example_proximate}
recipe_pm_fat <- preprocess_recipe(
  prep_smooth(w = 7, algorithm = "moving-average"),
  prep_snv(),
  prep_derivative(m = 1, w = 5, p = 11, algorithm = "nwp"),
  device = "proximate"
)

X_fat_prep <- process(X, recipe_pm_fat)
head(X_fat_prep[, 1:5])
```

## Example 2: ProxiScout workflow with detrending

ProxiScout instruments benefit from additional preprocessing steps:

```{r example_proxiscout_full}
recipe_ps_full <- preprocess_recipe(
  prep_resample(grid = "proxiscout"),
  prep_smooth(w = 7, p = 1, algorithm = "savitzky-golay"),
  prep_snv(),
  prep_detrend(p = 2),
  prep_derivative(m = 1, w = 5, p = 2, algorithm = "savitzky-golay"),
  device = "proxiscout"
)

X_ps_full <- process(X, recipe_ps_full)
dim(X_ps_full)
```

## Example 3: Minimal preprocessing

Sometimes less is more. A minimal recipe with only SNV:

```{r example_minimal}
recipe_minimal <- preprocess_recipe(prep_snv())
X_minimal <- process(X, recipe_minimal)
```

## Example 4: Wavelength band selection

Select a specific wavelength range for ProxiScout:

```{r example_band}
recipe_band <- preprocess_recipe(
  prep_wav_trim(band = c(1100, 1600)),
  prep_smooth(w = 5, p = 1, algorithm = "savitzky-golay"),
  prep_snv(),
  device = "proxiscout"
)

X_band <- process(X, recipe_band)
colnames(X_band)[c(1, ncol(X_band))]
```

# Best practices

## Order matters

Preprocessing steps affect each other. Common orderings:

1. **Smoothing → SNV → Derivative** (standard for noise reduction + normalization + enhancement)
2. **Resampling → Smoothing → Derivative → SNV** (device-specific resampling first)
3. **SNV → Detrend** (full de-trending procedure)

## Device-aware development

Always specify `device = "proximate"` or `device = "proxiscout"` when building recipes (except for SNV-only recipes). This ensures recipes are portable and the preprocessing is compatible with the target device.

## Reproducibility

Store recipes alongside calibration models to ensure preprocessing is applied identically during prediction. The `process()` function attaches the recipe as an attribute for downstream tracking.

## Parameter tuning

Preprocessing parameters are typically tuned during model development:

- Window sizes should be small relative to spectral features of interest
- Derivative orders depend on spectral complexity (first-order for baseline removal, second-order for sharper features)
- SNV is almost universally beneficial for multiplicative interference

# Summary

The preprocessing recipe system in `proximetricsR` provides a structured, reproducible approach to spectral preprocessing:

- **Constructors** define individual steps with validation
- **Recipes** compose steps into device-aware pipelines
- **Process** applies recipes consistently
- **Attributes** preserve recipe information for downstream use

This design enables seamless integration with calibration workflows and ensures preprocessing is applied consistently from model development through deployment on BUCHI devices.

```{r cleanup, include = FALSE}
options(old_options)
```
