---
title: "ProxiMate: Building applications"
author: 
 - name: Leonardo Ramirez-Lopez and Claudio Orellano
   email: ramirez-lopez.l@buchi.com
   affiliation: Data Science Department, BUCHILabortechnik AG, Flawil, Switzerland
date: today
clean: true
bibliography: ["proximetricsR.bib"]
biblio-style: "apalike"
link-citations: true
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-overflow: wrap
    smooth-scroll: true
    html-math-method: mathjax
vignette: >
  %\VignetteIndexEntry{ProxiMate: Building applications}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{quarto::html}
---

```{r}
#| label: setup
#| include: false
old_options <- options(digits = 3)
```


# Introduction

A Proximate application comprises the following files:

-   Calibration data file (.tsv)

-   Local data file (.tsv)

-   Project files (.prj)

-   Calibration model files (.cal)

-   Application file (.nax)

-   Report files (.rtf)

-   Application metadata file (.nad)

A detailed description of the above files and the structure of the
application files can be found in the vignette [Structure of the
application files](af-proximate-structure-of-the-applications.html)

# Build an application

Building an application file requires a series of steps that span from
importing the calibration data file(s) to writing down the files
containing the data of the predictive models for the response variables.
In the following subsections all these steps will be explained.

First of all, the `proximetricsR` and the `prospectr` libraries must be
installed and loaded.

```{r loadlib, results = 'hide', include = FALSE}
# Load package, or use devtools::load_all() if in development
if (!requireNamespace("proximetricsR", quietly = TRUE)) {
  devtools::load_all()
}
library("proximetricsR")
```

```{r loadlib2, eval = FALSE}
library("prospectr")
library("proximetricsR")
```

## Read the calibration data and prepare it

Here we use a two DEMO files containing spectral data of soybean meal samples
measured with two different BUCHI ProxiMate devices (up-view mode). These 
datasets are available at a public repository of BUCHI demo data.

```{r read, results = 'show'}
# data location
data_loc <- "https://raw.githubusercontent.com/buchi-labortechnik-ag/demo_data/main/data/"

# location of the tsv file containing the spectral data (of soybean meal)
my_file_1 <- "SoybeanMeal_file1.tsv"
my_file_2 <- "SoybeanMeal_file2.tsv"

my_file_1 <- paste0(data_loc, my_file_1)
my_file_2 <- paste0(data_loc, my_file_2)

# read the files
mdata_1 <- proximate_read_data(my_file_1)
mdata_2 <- proximate_read_data(my_file_2)


# explore the structure of the objects of the imported data sets
str(mdata_1, vec.len = 0, give.attr = FALSE)

str(mdata_2, vec.len = 0, give.attr = FALSE)
```

The spectral data is stored as a matrix inside the data object. It can
be accessed by:

```{r spc, results = 'hide'}
mdata_1$spc
mdata_2$spc
```

The wavelengths of the spectral data can be obtained from the column
names of the spectral matrix as follows:

```{r wavelengths, results = 'hide'}
original_wavs_1 <- as.numeric(colnames(mdata_1$spc))

original_wavs_2 <- as.numeric(colnames(mdata_2$spc))
```

The spectra loaded can be plotted using the `plot()` function:

```{r plotspectra, fig.cap = "Spectra loaded. Left: datastet 1; Right: Dataset 2",  fig.cap.style = "Image Caption", fig.align = "center", fig.width = 8, fig.height = 8, echo = TRUE, fig.retina = 0.85}
old_par <- par(no.readonly = TRUE)
par(mfrow = c(2, 1), mar = c(4, 4, 1, 4))

y_max <- max(c(max(mdata_1$spc), max(mdata_2$spc)))

matplot(
  x = original_wavs_1,
  y = t(mdata_1$spc),
  xlab = "Wavelengths, nm",
  ylab = "Absorbance",
  xlim = c(400, 1700),
  ylim = c(0, y_max),
  type = "l",
  lty = 1,
  col = rgb(1, 0.5, 0.5, 0.3)
)
grid()

matplot(
  x = original_wavs_2,
  y = t(mdata_2$spc),
  xlab = "Wavelengths, nm",
  ylab = "Absorbance",
  xlim = c(400, 1700),
  ylim = c(0, y_max),
  type = "l",
  lty = 1,
  col = rgb(0.9, 0.8, 0.5, 0.3)
)
grid()
```

## Merge multiple datasets

The two previous vectors (`original_wavs_1` and `original_wavs_2`) are different
as the wavelength positions of the two datasets are different which prevents 
direct merging of their spectral data. To solve this problem, the function 
`proximate_merge()` can be used:

```{r nwpmerge, results = 'hide'}
mdata <- proximate_merge(list(mdata_1, mdata_2))
```

The spectra in the resulting `mdata` object uses the wavelengths of the 
first dataset in the list of datasets as the reference wavelengths on which the 
spectral data of the remaining datasets will be resampled to. In this case, the
`proximate_merge()` function resampled both, `mdata_2$spc` and `mdata_3$spc` 
to the wavelengths of `mdata_1$spc`.


## Resample to constant resolution

One of the constrains to work with the the `proximetricsR` package is that
the spectra must be at a constant resolution. The spectra in this
example does not have a constant resolution, however, we can compute the
spectral values at a given set of equally spaced wavelengths (within the
same spectral range of the original set of wavelengths).

```{r resample}
# define the vector of new wavelengths (with constant resolution)
# (replace the psc matrix with the new resampled maxtrix)
# for working with proximetricsR we need to have our spectra in a spc matrix inside
# the data matrix
mdata$spc <- process(mdata$spc, prep_resample(c(900, 1700, 2)), device = "proximate")
```

The final wavelengths are then:
```{r finalwavs,  results = 'hide'}
final_wavs <- as.numeric(colnames(mdata$spc))
```

```{r plotspectra2, fig.cap = "Merged spectra.",  fig.cap.style = "Image Caption", fig.align = "center", fig.width = 8, fig.height = 5, echo = TRUE, fig.retina = 0.85}
par(old_par)
matplot(
  x = final_wavs,
  y = t(mdata$spc),
  xlab = "Wavelengths, nm",
  ylab = "Absorbance",
  type = "l",
  lty = 1,
  col = rgb(0.5, 0.5, 0.5, 0.3)
)
grid()
```



## Activate/deactivate rows for modeling

In case one or more rows need to be deactivated for modeling, the values of 
the `Check` column at these row indices need to be set to ``"False"`. Note that 
the deactivated rows are not removed from the dataset. For instance, if the rows 
`5`, `10` and `50` need to be deactivated as follows:  

```{r settoflase,  results = 'hide'}
# a vector with the row indices of the samples to be deactivated
deact <- c(5, 10, 50)
# deactivate
mdata$Check[deact] <- "False"
```

For this example, all the samples/rows in the final dataset will be used for 
modeling. For that, it is necessary that all the values in the `Check` column 
need to be set to `"True"`.
```{r settotrue,  results = 'hide'}
mdata$Check <- "True"
```


## Get the names of the response variables

A trick to get the names of all the response variables in the data

```{r properties, results = 'hide'}
# this gets the names of all the variables in the data
names(mdata)

# this returns the column names of all responses
y_names <- extract_property_names(mdata)

# the indices of all the response variables
ys_indices <- which(colnames(mdata) %in% y_names)

# samples with reference values
colSums(!is.na(mdata[, y_names]))
```

From the `r length(y_names)` properties, the examples that will follow will only 
take into account 2 of them (the first two): `r paste0(y_names[1:2], collapse = " and ")`.


```{r properties3, results = 'hide'}
# get the names of the response variables
y_names <- y_names[1:2]
```

## Calibrating models using `calibrate_models()`

The `calibrate_models()` allows users to automatically build multiple
calibration models with the optimization of the pre-processing methods.
This function uses internally another function named `calibrate()` (type 
`?calibrate` in your R console) for calibrating individual models. 
The arguments of `calibrate_models()` resemble those of `calibrate()`. 

Apart from the input calibration data, the `calibrate_models()` function has 
other 5 main arguments: 

- `formulas`: specifies the target formulas of the target models that 
need to be built. 

- `metadata_list`: it provides the specifications for the metadata of each model 
in formulas (e.g. the units of the properties). 

- `preprocess_recipes`: specifies the sequences of preprocessing methods 
(preprocessing recipes) to be tested. 

- `method`: indicates the type of regression method to use along 
with its parameters (e.g. pls or xls). 

- `control`: indicates how some aspects of the calibration process must be 
conducted (e.g. cross-validation and outlier detection, etc). 



### `formulas`: defining what is to be modeled

Here the predictive functions to be calibrated are defined (e.g. protein as 
function of spectra) by using the generic way of writing function formulas in R. 
To do so, the name of the response variable is given first, followed by the 
tilde operator (`~`) and then followed by the name of the matrix of spectral 
predictors. The names of the response variables need to match their names in the 
respective columns. As the spectral variables are stored under a matrix in the 
dataset (typically named `spc`), the user can directly put the name of the 
matrix instead of writing the names of the wavelengths. For example:

```{r formula1, results = 'hide'}
f_simple <- Protein ~ spc

# or

f_long <- as.formula(paste0("Protein ~ ", paste0(final_wavs, collapse = " + ")))
```

In the previous example, both `f_simple` and `f_long` are equivalent. 

In `proximetricsR`, the user can also evaluate multiple functions (for different 
response variables) at once. For that it is necessary to create a list of 
functions:

```{r formulas1, results = 'hide'}
app_formulas <- list(Moisture ~ spc, Protein ~ spc)
```

Alternatively, if we already have the property names, we can do the following to 
easily build the model formulas:

```{r formulas2, results = 'hide'}
app_formulas <- lapply(paste0(y_names, " ~ spc"), FUN = formula)
app_formulas
```


### `metadata_list`: specify the property metadata

Apart from defining the formulas, the user can also indicate some metadata for 
each model formula (e.g. the units and the decimal places to be shown by the 
sensor when displaying a predicted value). In this exercise, this is done via:

```{r metadata, results = 'hide'}
# Define the metadata of each model (in the same order they are in app_formulas)
models_metadata <- list(
  # for the first property
  add_model_metadata(decimal_places = 2, unit = "%"),
  # for the second property
  add_model_metadata(decimal_places = 2, unit = "%")
)
```

Note that the metadata of each model must be specified in the same order of 
response variables as in `app_formulas`. 


### `preprocess_recipes`: create pre-processing recipes

One of the advantages of using this package is that it uses the concept
of pre-processing recipe. A such recipe is basically the instructions on
how the different pre-processing steps must be executed and what
parameters must be used in each step. The recipe must be built/ordered
in the same sequence the pre-processing steps must be applied. The
following lines of code provide an example on how a recipe is built:

```{r recipe, results = 'show'}
# here a recipe with the following pre-processing steps is created:
# - spline/resampling: with spectral range between 900 and 1700 in steps of 4
# - StandardNormalVariate
# - First derivative: with a gap parameter of 5 and a smoothing factor of 9
my_precipe_1 <- preprocess_recipe(
  prep_resample(c(900, 1700, 4)),
  prep_snv(),
  prep_derivative(m = 1, w = 5, p = 9, algorithm = "nwp"),
  device = "proximate"
)

my_precipe_1$preprocessing_order
```

Similarly, other recipes can be configured:

```{r recipe2, results = 'show'}
# here a recipe with the following pre-processing steps is created:
# - spline/resampling: with spectral range between 900 and 1700 in steps of 5
# - StandardNormalVariate
# - First derivative: with a gap parameter of 7 and a smoothing factor of 11
my_precipe_2 <- preprocess_recipe(
  prep_resample(c(900, 1700, 4)),
  prep_snv(),
  prep_derivative(m = 2, w = 7, p = 11, algorithm = "nwp"),
  device = "proximate"
)

my_precipe_2$preprocessing_order

# here a recipe with the following pre-processing steps is created:
# - spline/resampling: with spectral range between 900 and 1700 in steps of 4
# - StandardNormalVariate
# - Second derivative: with a gap parameter of 9 and a smoothing factor of 13
my_precipe_3 <- preprocess_recipe(
  prep_resample(c(900, 1700, 4)),
  prep_snv(),
  prep_derivative(m = 2, w = 9, p = 13, algorithm = "nwp"),
  device = "proximate"
)

my_precipe_3$preprocessing_order

# here a recipe with the following pre-processing steps is created:
# - spline/resampling: with spectral range between 900 and 1700 in steps of 8
# - StandardNormalVariate
# - Second derivative: with a gap parameter of 5 and a smoothing factor of 5
my_precipe_4 <- preprocess_recipe(
  prep_resample(c(900, 1700, 5)),
  prep_snv(),
  prep_derivative(m = 2, w = 5, p = 5, algorithm = "nwp"),
  device = "proximate"
)

my_precipe_4$preprocessing_order
```

When calibrating spectral models, usually different pre-processing recipes are 
tested. The best recipe is chosen based (at least partially) on the predictive
performance. To make the testing procedure simple, `proximetricsR` allows to test
multiple recipes at once. For this, it is necessary to group the recipes into a 
list, as follows: 

```{r recipelist, results = 'show'}
my_precipes <- list(my_precipe_1, my_precipe_2, my_precipe_3, my_precipe_4)
```

The `my_precipes` list will be later use for modeling. 


### `method`: how to fit the spectral models

We need to define what regression method is to be used for the calibration of 
the model(s). This can be done as follows:

```{r rmethod, results = 'show'}
# here we use the modified pls regression method (equivalent to the one
# implemented in NIRWise PLUS)
# We use a maximum 11 components
my_pls_method <- fit_plsr(ncomp = 11, type = "nwp")
my_pls_method
```

In addition to the `fit_plsr()`, the `fit_xlsr()` is also available (see the 
vignette ___Mathematical overview of regression algorithms___ for details on 
both methods).

```{r rmethodxls, results = 'show'}
# here we use the modified pls regression method (equivalent to the one
# implemented in NIRWise PLUS)
# We use a maximum 11 components
my_xls_method <- fit_xlsr(ncomp = 11)
my_xls_method
```


### `control`: Setting up the calibration parameters

To build calibration models for each response variable we need to
define some generic aspects that control the way in which calibrations
are build (e.g. how validation is to be conducted, how to handle
outliers, etc). To define such overall aspects, the
`calibration_control()` function can be used. The following code
exemplifies how to use that function to define the cross-validation type
and the outlier detection:

```{r calcontrol, results = 'show'}
# Control some aspects of how the calibration is built and optimized
# We use k-fold cross-validation with selection of folds based on the order of
# the samples in the data table ("sequential")
# We also specify the number of times a model must be re-fitted after outlier
# removal, e.g. 0 means no re-fiting i.e. no outlier removal; 1 means a model is
# built, then it is used to identify and remove outliers and finally a the final
# model is refitted; a value of 5 would mean that the model is refitted 4 times
# for 4 outlier removal iterations.
my_control <- calibration_control(
  validation_type = "kfold",
  number = 4,
  folds = "sequential",
  remove_outliers = 1 # the number of iterations of outlier removal
)
```

Go to the help page of the `calibration_control()` to obtain more information 
on all the calibration aspects that can be set with this function. 


## Calibrate the spectral models

Once the pre-processing recipes, the calibration parameters and
the functions to be fitted, the code for calibrating the different models is
rather simple. For example:

```{r calibration, results =  'hide'}
optimized_app <- calibrate_models(
  formulas = app_formulas,
  data = mdata,
  preprocess_recipes = my_precipes,
  methods = list(my_pls_method, my_xls_method),
  control = my_control,
  metadata_list = models_metadata,
  save_all = TRUE
)
```


```{r showoptimize2, eval = FALSE}
optimized_app
```
The above object shows you a table with the validation results and the suggested
pre-processing methods per model:

```{r showoptimize1, eval = TRUE, echo = FALSE}
print(optimized_app, separator = " >\n")
```

The above object prints a summary of the model optimization results.
The best models are marked with an asterisk (`*`). The preprocessing recipes are 
numbered according to their order in the list passed to the `preprocess_recipes` 
argument. 

### Overview of the best models found

These best models are stored in the resulting object under 
`...$final_models`:

```{r finalmodels1, eval =  FALSE}
optimized_app$final_models
```


```{r finalmodels2, eval =  TRUE}
names(optimized_app$final_models)
```

For the above final models, `r names(optimized_app$final_models)[1]` and 
`r names(optimized_app$final_models)[2]`, the best preprocessing recipes are the
ones in `my_precipes` with indices `r optimized_app$results_grid$recipe[optimized_app$results_grid$selection][1]` 
and `r optimized_app$results_grid$recipe[optimized_app$results_grid$selection][2]`
respectively. This is, for `r names(optimized_app$final_models)[1]`:

```{r prepro1, echo = FALSE, eval =  TRUE, comment = NA}
print(optimized_app$final_models[[1]]$preprocess, separator = " >\n")
```

and for `r names(optimized_app$final_models)[2]`:
```{r prepro2, echo = FALSE, eval =  TRUE, comment = NA}
print(optimized_app$final_models[[2]]$preprocess, separator = " >\n")
```

These recipes can be accessed by:
```{r prepro3, eval =  FALSE}
optimized_app$final_models$`Protein ~ spc`
# and
optimized_app$final_models$`Moisture ~ spc`
```

A comprehensive set of interactive plots describing each model can be obtained 
by:
```{r modelplots, eval =  FALSE}
plot(optimized_app$final_models$`Protein ~ spc`, selection = "all")
# and
plot(optimized_app$final_models$`Moisture ~ spc`, selection = "all")
```

Applying the `plot()` function on a model produces a HTML file which is 
immediately loaded and shown in a browser.


The summary table of the optimization can be accessed by:
```{r summaryopt, eval =  FALSE}
optimized_app$results_grid
```

The `results_grid` table comprises the following fields:

- `formula`: the formulas used to calibrate the spectral models of the properties.

- `recipe`: the index of the preprocessing recipe in the list of preprocessings 
passed to the `calibrate_models()` function.

- `ncomp`: the optimal number of components found.

- `rsq`: the squared correlation coefficient. 

- `rmse`: the rrot mean squared error.

- `largest_residual`: the value of the largest residual found.

- `outliers`: the number of outliers detected and removed.

- `selection`: a logical indicating the final models found for each property.


The `TRUE` values in the column `selection` indicate the final models found. 
This column can be used to filter the table to show only the description of the
final models:
```{r summaryopt2, eval =  TRUE}
optimized_app$results_grid[optimized_app$results_grid$selection, ]
```

### Checking all the models tested

It is possible to use as final models others than the ones suggested/found by 
the `calibrate_models()` function. In that case, these models can be extracted 
from the list of all models tested, as follows:

```{r othermodels, eval =  FALSE}
optimized_app$all_models
```

The models in `optimized_app$all_models` are indexed according to the order 
given in the `optimized_app$results_grid` table. For example, to manually extract
the models 1 (Protein using preprocessing recipe 1) and 6 (Protein with 
preprocessing recipe 6):

```{r othermodels2, eval =  FALSE}
my_other_models <- optimized_app$all_models[c(1, 6)]
```


## Predicting the properties in unseen samples

In this example we will load another (and independent) file containing 
measurements of soybean meal samples: 

```{r preds, eval =  TRUE, results = 'hide'}
my_pred_file <- "SoybeanMeal_file3.tsv"
my_pred_file <- paste0(data_loc, my_pred_file)

pdata <- proximate_read_data(my_pred_file)
original_wavs_pred <- as.numeric(colnames(pdata$spc))
```

```{r plotspectra3, fig.cap = "Merged spectra.",  fig.cap.style = "Image Caption", fig.align = "center", fig.width = 8, fig.height = 5, echo = TRUE, fig.retina = 0.85}
matplot(
  x = original_wavs_pred,
  y = t(pdata$spc),
  xlab = "Wavelengths, nm",
  ylab = "Absorbance",
  type = "l",
  lty = 1,
  col = rgb(0.3, 0.9, 0.5, 0.5)
)
grid()
```

Predicting for the samples in `pdata` the response values, for which the models 
in `optimized_app` were calibrated for, is straightforward. For that, the new 
data loaded must contain the predictor variable(s). In this case, our predictor 
is `spc` which contains all the necessary spectral variables used for 
calibrating the models. We then can use this data along with the `predict` 
function and the object containing the models as follows:

```{r preds2, eval =  TRUE, results = 'hide'}
preds <- predict(optimized_app, newdata = pdata)

# matrix of predictions
preds$predictions

# info on the models used
preds$model_information
```

The `preds`object generated by `predict()` comprises `r `length(names(preds))`
components: `r names(preds)[1]` and `r names(preds)[2]`. The first one contains 
a matrix of predictions while the second one contains a list with descriptions 
of the models that were used. 

If the reference values are available in the prediction data (as in the example),
we can then compare/validate such predictions. For example:

```{r preds3, fig.cap = "Validation plots for the preditions of the optimal models found",  fig.cap.style = "Image Caption", fig.align = "center", fig.width = 7, fig.height = 3.5, echo = TRUE, fig.retina = 0.85}
old_par <- par(no.readonly = TRUE)
par(mfrow = c(1, 2), mar = c(4, 4, 1, 1))

# Moisture plot
plot(
  x = preds$predictions[, "Moisture"],
  y = pdata$Moisture,
  xlab = "Predicted moisture, %",
  ylab = "Measured moisture, %",
  xlim = c(10.5, 12.5),
  ylim = c(10.5, 12.5),
  pch = 16,
  cex = 1.5,
  col = "grey"
)
grid()
abline(0, 1, col = "red")

# Protein plot
plot(
  x = preds$predictions[, "Protein"],
  y = pdata$Protein,
  xlab = "Predicted protein, %",
  ylab = "Measured protein, %",
  xlim = c(44, 48),
  ylim = c(44, 48),
  pch = 16,
  cex = 1.5,
  col = "grey"
)
grid()
abline(0, 1, col = "red")
par(old_par)
```

The squared correlation coefficient and the root mean squared error (rmse) can 
be computed by:

```{r preds4, eval =  FALSE, results = 'hide'}
# squared correlations
cor(x = preds$predictions[, "Moisture"], y = pdata$Moisture)^2
cor(x = preds$predictions[, "Protein"], y = pdata$Protein)^2

# rmses
mean((preds$predictions[, "Moisture"] - pdata$Moisture)^2)^0.5
mean((x <- preds$predictions[, "Protein"] - pdata$Protein)^2)^0.5
```

## Writting down an application file

An application is basically a collection of calibration models with some
additional parameters (see the ___Structure of the application files___
vignette). This application can be written into a file that
can be imported into the ProxiMate sensor to then be used to predict
response variables from the spectral measurements made with that sensor.
The `proximetricsR` package also offers a function to write down this
application file. For this, the models that the user wants to include in
the application need to be placed together into a `list` object. For
example:

```{r mymodelslist, eval = TRUE, results = 'hide'}
my_models <- optimized_app$final_models

# add some important metadata to the application/model list
my_models <- add_application_metadata(
  my_models,
  view = "Up",
  name = "my_application",
  description = "created with proximetricsR"
)
```

The order in which the objects are declared in the list will be the same
order in which their corresponding response variables will be shown in
the ProxiMate sensor.

The application for the calibration models inside `my_models` can be
finally be written as follows:

```{r appfile, eval = FALSE, results = 'hide'}
proximate_write_nax(
  path = getwd(),
  object = my_models
)
```


# Other functionality

## Calibrate single models with `calibrate`


```{r protein, eval = FALSE, results = 'hide'}
my_p_recipe <- preprocess_recipe(
  prep_resample(c(900, 1700, 2)),
  prep_snv(),
  prep_derivative(m = 1, w = 5, p = 3, algorithm = "nwp"),
  device = "proximate"
)

protein_model <- calibrate(
  Protein ~ spc,
  data = mdata,
  preprocess = my_p_recipe, # here we can use only one recipe
  method = fit_xlsr(5, type = "nwp"),
  control = my_control
)
protein_model

# add the units and the number of decimal places to the model
protein_model <- add_model_metadata(
  protein_model,
  unit = "%",
  decimal_places = 2
)

plot(protein_model, selection = "all")
```

The fitted model can also be used to make predictions on new data. Just
for demonstration purposes, here we will use the `mdata` data set (the
one used for model building) to show how predictions can be obtain from
the model applied on spectra:

```{r predictions, results = 'hide', eval = FALSE}
# add the units and the number of decimal places to the model
my_protein_predictions <- predict(protein_model, newdata = pdata$spc)
my_protein_predictions$model_information
my_protein_predictions$predictions
```

```{r writenax, results = 'hide', eval = FALSE}
# create a list with all the models separately calibrated
# (in this case we only have one)
a_model <- list(protein_model)

a_model <- add_application_metadata(
  a_model,
  view = "Up",
  name = "a_protein_application",
  description = "created with proximetricsR"
)

proximate_write_nax(
  path = getwd(),
  object = a_model
)
```

The calibration of a single model can also be useful when there is a model that
we might want to replace  in the list of models suggested by `calibrate_models()`.
For example:
```{r replacemodel, results = 'hide', eval = FALSE}
my_models <- optimized_app$final_models
my_models$`Protein ~ spc` <- protein_model
```

## Write model-related files (`tsv`, `cal`, `prj` and `rtf`)

The calibration (.cal), project (.prj) and report (.rtf) files native of the
NIRWise PLUS software can be written down using the `proximate_write_model()`
function. To be able to use such files, it is necessary to have local access to
the data (.tsv) file containing the calibration data used to fit the models. For example:

```{r writetsv, eval = FALSE, results = 'hide'}
# let's write down the calibration data into a tsv file
# define a name for the tsv to be written
tsv_name <- "merged_cal_data.tsv"

# write the tsv
proximate_write_data(
  x = mdata,
  file = tsv_name,
  id = mdata$ID,
  spc = "spc",
  spc_round = 8,
  barcode = "",
  properties = y_names,
  note = "merged_file_1_and_file_2",
  recipe = "SoybeanMeal up-view",
  created = mdata$Date,
  snr = mdata$SNR
)
```

```{r prjcal, eval = FALSE, results = 'hide'}
proximate_write_model(
  object = list(protein_model),
  path = getwd(),
  tsv_paths = tsv_name,
  application_name = "Untitled",
  cal = TRUE,
  prj = TRUE,
  rtf = TRUE
)
```

Note that project files can be opened in NIRWise PLUS. So, if you have
access to that software you could try to open the .prj file written with
the above code.

```{r cleanup, include = FALSE}
options(old_options)
```
