---
title: "ProxiScout: Building applications"
author:
 - name: Leonardo Ramirez-Lopez and Claudio Orellano
   email: ramirez-lopez.l@buchi.com
   affiliation: Data Science Department, BUCHILabortechnik AG, Flawil, Switzerland
date: today
clean: true
bibliography: ["proximetricsR.bib"]
biblio-style: "apalike"
link-citations: true
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-overflow: wrap
    smooth-scroll: true
    html-math-method: mathjax
vignette: >
  %\VignetteIndexEntry{ProxiScout: Building applications}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{quarto::html}
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  eval = requireNamespace("proximetricsR", quietly = TRUE)
)
# Disable ANSI colours for vignette rendering
options(cli.num_colors = 1)
Sys.setenv("RSTUDIO" = "")
Sys.setenv("POSITRON" = "")
old_options <- options(digits = 3)
```

# Introduction

ProxiScout applications consist of calibrated predictive models packaged in a device-compatible format. Unlike ProxiMate's file-based structure (.nax, .cal, .prj), ProxiScout applications use JSON-based serialization synchronized through the NeoSpectra Portal.

A typical ProxiScout workflow involves:

1. Loading and preparing spectral data  

2. Building and validating calibration models  

3. Preprocessing configuration matching device algorithms  

4. Exporting models in ProxiScout-compatible format  

See the ProxiScout Structure vignette for detailed information on application file formats and metadata.

# Setup


```{r}
#| label: loadlib
#| results: 'hide'
#| include: False
library("proximetricsR")
```


```{r loadlib2, eval = FALSE}
library("proximetricsR")
```

# Workflow overview

## Prepare spectral data

Load your spectral data. Data should include: 

- Spectral matrix (samples × wavenumbers)  

- Reference property values (calibration targets)  

- Sample metadata (optional)  

Here, a demo dataset from the `prospectr` package [@prospectr2026]. This is used in these examples as it covers all the spectral regions measured by ProxiScout devices:

```{r data}
data("NIRsoil", package = "prospectr")

# NIRsoil is in nanometers and absorbance
mdata <- data.frame(
  sampleName = rownames(NIRsoil),
  Ciso = NIRsoil$Ciso,
  Nt = NIRsoil$Nt,
  CEC = NIRsoil$CEC
)

# ProxiScout data comes in reflectance in percentages, so we convert to
# reflectance and scale to 0-100 range
mdata$spc <- 100 * 1 / 10^(NIRsoil$spc)

# Get wavelengths and convert from nm to wavenumbers (cm^-1)
wav_nm <- as.numeric(colnames(mdata$spc))
wav_cm <- 10000000 / wav_nm

# Update column names to wavenumbers for ProxiScout compatibility
colnames(mdata$spc) <- wav_cm

head(colnames(mdata$spc))
```

```{r}
class(mdata) <- c("proxiscout_data", class(mdata))
```


ProxiScout works with wavenumbers (cm⁻¹) in the NeoSpectra range (~3922-7407 cm⁻¹). The NIRsoil data spans ~400-2500 nm (4000-25000 cm⁻¹), but resampling to the ProxiScout grid will retain only overlapping wavenumbers.

## Define preprocessing recipe

ProxiScout supports device-specific preprocessing via the NeoSpectra wavenumber grid. Define your preprocessing sequence:

```{r recipe}
recipe_01 <- preprocess_recipe(
  prep_resample(grid = "proxiscout"), # necessary for almost all ProxiScout recipe
  prep_derivative(m = 2, w = 9, p = 2, algorithm = "savitzky-golay"),
  prep_snv(),
  device = "proxiscout"
)

recipe_01
```

Key points:
- `prep_resample(grid = "proxiscout")` resamples to the standard NeoSpectra wavenumber grid, retaining only overlapping wavenumbers 

- Smoothing uses Savitzky-Golay (not moving-average)  

- Derivatives support Savitzky-Golay or gap-segment algorithms  

- Additional steps like `prep_detrend()`, `prep_transform()`, and `prep_wav_trim()` are available  

- Data is already in absorbance, so no conversion needed

## Build calibration model

### Example 1: Build a single model 

Apply preprocessing and build the model:

```{r calibrate, results = 'hide'}
model_c <- calibrate(
  Ciso ~ spc,
  data = mdata,
  preprocess = recipe_01,
  method = fit_plsr(ncomp = 12, type = "modified"),
  control = calibration_control(
    validation_type = "kfold",
    number = 5,
    seed = 42
  ),
  verbose = FALSE
)
```

Check model performance:

```{r evaluate, results = 'hide'}
model_c
```

```{r}
#| eval: false
plot(model_c)
```

#### Serialize the model for deployment 

The model can be exported to ProxiScout format for deployment. Here we can see
how that model is serialized:
```{r}
#| label: serialized
#| results: 'hide'
#| eval: false
my_serialised_model_c <- proxiscout_write_model(model_c, file = NULL)
my_serialised_model_c
```

To write the json file:

```{r}
#| eval: false
proxiscout_write_model(model_c, file = "my_model_c.json")
```


### Example 2: Build multiple models at once testing different pre-processings

To test multiple preprocessing recipes and build multiple models at once, use the `calibrate_models()` function. This function allows you to specify a list of formulas, preprocessing recipes, and modeling methods to systematically evaluate different combinations.

Here is an example of how to build multiple models at once:

First, build the list of formulas representing the calibration models we need
to build:
```{r}
my_formulas <- list(Ciso ~ spc, CEC ~ spc)
```

Now let's define multiple preprocessing recipes to test. For example, we can compare the performance of a simple derivative-based recipe with a more complex one that includes wavenumber trimming:

```{r}
#| label: multi_calibrate1
#| results: 'hide'
recipe_02 <- preprocess_recipe(
  prep_resample(grid = "proxiscout"),
  prep_derivative(m = 1, w = 7, p = 2, algorithm = "savitzky-golay"),
  device = "proxiscout"
)

recipe_03 <- preprocess_recipe(
  prep_resample(grid = "proxiscout"),
  prep_wav_trim(
    band = c(4000, 7000),
    trim_constant_edges = TRUE
  ),
  prep_derivative(m = 1, w = 7, p = 2, algorithm = "savitzky-golay"),
  device = "proxiscout"
)

my_recipes <- list(recipe_01, recipe_02, recipe_03)
```


Define the multiple model fitting methods:

```{r}
my_fitting <- list(
  fit_plsr(ncomp = 12, type = "modified"),
  fit_plsr(ncomp = 10, type = "standard")
)
```

Then, define the how to control the validations. This object will be used for all the models we are going to build:

```{r}
my_control <- calibration_control(
  validation_type = "kfold",
  number = 5,
  remove_outliers = 1,
  seed = 42
)
```

Finally, build all the models at once:

```{r}
#| label: multi_calibrate2
#| results: 'hide'
multiple_models <- calibrate_models(
  formulas = my_formulas,
  data = mdata,
  preprocess_recipes = my_recipes,
  methods = my_fitting,
  control = my_control,
  save_all = TRUE
)
```


```{r}
#| eval: false
multiple_models
```

```{r}
#| eval: true
#| echo: false
df <- multiple_models$results_grid
num_cols <- sapply(df, is.numeric)
df[num_cols] <- lapply(df[num_cols], round, 2)
multiple_models$results_grid <- df
print(multiple_models)
```

#### Serialize the multiple models for deployment 


```{r serialized2}
#| eval: false
proxiscout_write_model(
  multiple_models$final_models$`Ciso ~ spc`,
  file = "my_model_c2.json"
)

proxiscout_write_model(
  multiple_models$final_models$`CEC ~ spc`,
  file = "my_model_cec.json"
)
```

## Export for ProxiScout

Once satisfied with model performance, export to ProxiScout format. Refer to the Structure vignette for export details.

# Device-specific considerations

**NeoSpectra wavenumber grid:** ProxiScout instruments measure at fixed wavenumber positions. Always resample to `"proxiscout"` grid when building models for deployment.

**Algorithm selection:** Choose from:   

- Smoothing: Savitzky-Golay only   

- Derivatives: Savitzky-Golay or gap-segment  


- Additional: detrending, reflectance/absorbance conversion

**Advanced preprocessing:** ProxiScout's broader algorithm support allows more flexible preprocessing pipelines compared to ProxiMate.

```{r cleanup, include = FALSE}
options(old_options)
```  

# References {-}


